RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-04 doc/paper.tex
Do strane 16

\subsection{Parameter tuning}
The \textsc{Rils}-\textsc{Rols} algorithms, apart of the most other methods for SR, employs just two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is set to 0.001(?), according to the preliminary experimental evaluation. The second parameter is tuned monitoring the two errors the algorithm produced on the training set; we choose $tolerance_{error}=?$ for which a largest average $R^2$ and the smallest average RMSE error over all instances from the training set are produced. Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on real--world benchark sets}
In this section, we evaluate the performance of \textsc{Rils-Rols} algorithm with 15 other competitors from the literature that solve the problem of Symbolic regression. The results are reported in terms of three numerical Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}, where two real-world benchmark sets, \textsc{Feynman} and \textsc{Storgatz} are considered, utilizing  different values of the noise levels: 0.0 (no noise), 0.001 (a low-level noise), and 0.01 (a high-level noise), respectively. Concerning the added noise, the White Gaussian noise was added to the target ($y$-axis values) as a fraction of the signal RMS value, also applied in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$, we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{N} \sum {y_i^2}}\right). $$

Each table consist of three blocks. The first block reports the algorithm whose performances are reported. The second block reports, in two columns,  percentages of instances where the model which matches to the exact one has been found by respective algorithms on both, \textsc{Feynman} and \textsc{Strogatz}, benchmark sets. The third column of this block provides the overall summary percentage of success over all instances. The third block displays the results of the respective algorithm in the presence of $R^2$ error which can indicate on  accuracy, i.e. in our cases when the algorithm  at least reached the accuracy ($R^2$ score) of 0.999.  Note that the termination of our algorithm may not necessary be triggered by the error equal to 1. These results are of interest as it may point out the existence of an over-fitting. This block consists of three rows that indicate the percentage of instances where respective algorithm is able to reach the desired accuracy (fulfilling condition $R^2 \geq 0.999$ before a termination due exceeding the time limit) on the both real--world benchmark sets separately, and the overall percentage of success rate (over all considered instances) w.r.t. the same criterion of success.

@@

\begin{itemize}
	\item Concerning the exact results in case no noise is utilized in the instance problems, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, begin successful in 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and just 27.14\% of all instances of the benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered benchmark sets. 
	\item   Concerning the exact results when the target noise is presented with a level of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model in more than 42\% cases considering the problems from the both benchmark sets. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the increase in the target noise level, as its performances rapidly drop down -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case the \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} in this order of success, whose percentages of success rate range from $\approx$15--17\%. 
	\item  Concerning the exact results when the presence of noise is significant, that is with a level of 0.01,  our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model for  34.77\% of the all problem instances. The second best is \textsc{Gplearn} which percentage of success is just 12.75\%, followed by \textsc{Dsr} which is successful for just 9.23\% of the all problem instances.

>>>>

\subsection{Parameter tuning}
The \textsc{Rils}-\textsc{Rols} algorithm employs just two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is fixed to 0.001(?), according to the preliminary experimental evaluation we conducted. The second parameter is tuned by monitoring the two errors the algorithm produced on a randomly chosen training set of xx problem instances; among the following values \{ x, xx, ... \} we chose $tolerance_{error}=?$ for which the largest average $R^2$ and the smallest average RMSE error over all problem instances from the training set are produced. %Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on real--world benchark sets}
In this section, we evaluate the performance of \textsc{Rils-Rols} algorithm and compare it to the 14 other competitors from the literature. The numerical results are reported in terms of the three Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}. The results for the two ground-truth benchmark sets, \textsc{Feynman} and \textsc{Storgatz}, are given w.r.t.\   the following values of the noise levels of 0.0 (no noise), 0.001 (a low-level noise), and 0.01 (a high-level noise), respectively. In details, the white Gaussian noise was added to the target ($y$-axis) values as a fraction of the signal RMS value, also applied in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$, we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{n} \sum {y_i^2}}\right),$$
where $n$ relates to the number of items in the input data.

Each table consists of three blocks. The first block reports the algorithm whose performances are reported (by a row). The second block reports, with two columns,  percentages of the instances where the model which fits to the exact one has been found by respective algorithms on \textsc{Feynman} and \textsc{Strogatz} benchmark set, respectively. The third column of this block provides the overall summary exact percentage of success over all instances from the both benchmark sets. The third block displays the results of the respective algorithm in terms of the $R^2$ score which indicate on  accuracy, i.e. in our cases when the algorithm  at least reached the accuracy ($R^2$ score) of 0.999.  Note that the termination of our algorithm may not always be triggered by reaching an appropriate $R^2$ score of 0.999, but also due to the reached time limit. These results are of interest as when compared to corresponding exact percentage it may point out the existence of an over-fitting. This block consists of three rows that indicate the percentage of instances where respective algorithm is able to reach the desired accuracy (fulfilling the condition $R^2 \geq 0.999$ upon termination) on the both real--world benchmark sets separately, and the overall percentage of success rate (over all considered instances) w.r.t. the same criterion of reaching the desired accuracy.

@@

\begin{itemize}
	\item \fxnote{Marko: Came up to here!}Concerning the exact results in case no noise is utilized in the instance problems, the best performing algorithm is our \textsc{Rils}--\textsc{Rols}, able to find the right model on 57.84\% problems instances of the benchmark set \textsc{Feynman}; the same algorithm was even more impressive in solving problem instances from benchmark set \textsc{Strogatz}, begin successful in 83.57\% cases. The second best performing  approach is \textsc{AI-Feynman}, successful in solving 55.78\% and just 27.14\% of all instances of the benchmark set \textsc{Feynman} and \textsc{Strogatz}, respectively. All other approaches are performing significantly worse, and none of them is able to solve more than 30\% instances of any of the two considered benchmark sets. 
	\item   Concerning the exact results when the target noise is presented with a level of 0.001, our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model in more than 42\% cases considering the problems from the both benchmark sets. Interestingly, \textsc{AI-Feynman} is quite sensible w.r.t.\ the increase in the target noise level, as its performances rapidly drop down -- it is successful on just 2.81\% of the all problem instances. Better from \textsc{AI-Feynman} are in this case the \textsc{Dsr}, \textsc{Gplearn}, and \textsc{Afp-Fe} in this order of success, whose percentages of success rate range from $\approx$15--17\%. 
	\item  Concerning the exact results when the presence of noise is significant, that is with a level of 0.01,  our \textsc{Rils}--\textsc{Rols} is still performing reasonably well, having found an exact model for  34.77\% of the all problem instances. The second best is \textsc{Gplearn} which percentage of success is just 12.75\%, followed by \textsc{Dsr} which is successful for just 9.23\% of the all problem instances.
